---
title: "07-dada2-HCdata"
author: "Joe Duprey"
date: "3/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Packages
load dada2

```{r packages}
library(dada2); packageVersion("dada2")
```
lets find the fastqs

```{r fastq path}
path <- "../data/demultiplexed" # directory containing fast qs
list.files(path)
```
Now we read in the names of the fastq files, and perform some string manipulation to get matched lists of the forward and reverse fastq files.
```{r}
# Forward and reverse fastq filenames have format: SAMPLENAME_R1_001.fastq and SAMPLENAME_R2_001.fastq
fnFs <- sort(list.files(path, pattern="Fwd", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="Rev", full.names = TRUE))
# Extract sample names, assuming filenames have format: SAMPLENAME_XXX.fastq
sample.names <- sapply((basename(fnFs)), `[`, 1)
```
## Inspect read quality profiles
We start by visualizing the quality profiles of the forward reads:
```{r}
plotQualityProfile(fnFs[1:2])
```
Now we can view the quality of the reverse reads.
```{r}
plotQualityProfile(fnRs[1:2])
```
## Trimming!
Assign the filenames for the filtered fastq files
```{r}
# Place filtered files in filtered/ subdirectory
filtFs <- file.path(path, "filtered", paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(path, "filtered", paste0(sample.names, "_R_filt.fastq.gz"))
names(filtFs) <- sample.names
names(filtRs) <- sample.names
```
We’ll use standard filtering parameters: maxN=0 (DADA2 requires no Ns), truncQ=2, rm.phix=TRUE and maxEE=2. The maxEE parameter sets the maximum number of “expected errors” allowed in a read, which is a better filter than simply averaging quality scores.
```{r}
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(220,220),
              maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
              compress=TRUE, multithread=TRUE) # On Windows set multithread=FALSE
head(out)
```
## Learn the Error Rates
"The DADA2 algorithm makes use of a parametric error model (err) and every amplicon dataset has a different set of error rates. The learnErrors method learns this error model from the data, by alternating estimation of the error rates and inference of sample composition until they converge on a jointly consistent solution. As in many machine-learning problems, the algorithm must begin with an initial guess, for which the maximum possible error rates in this data are used (the error rates if only the most abundant sequence is correct and all the rest are errors)." -dada2 tutorial
```{r}
errF <- learnErrors(fnFs, multithread=TRUE)
```

